Pyg4ometry: a python library for the creation of Monte Carlo radiation transport physical geometries

### 2019-11-30 public/paper/pyg4ometry.tex
Small changes
Updated all papers

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

There are existing codes that have functionality similar to pyg4ometry. Tools exist to convert CAD files to GDML, e.g. GUIMesh \cite{GUIMesh_Pinto}.
CAD software can export data to STL. The existent set of software does not provide a complete set of tools to efficiently create complex geometries.

@@

\subsection{Geant Detector Mark-up Language} 
To exchange geometry descriptions Geant Detector Mark-up Language (GDML) was developed. Geant4 and ROOT can read and write GDML format and it is commonly
used as an exchange format for Geant4 geometries. 
\begin{table}[hbt!]
\centering
\begin{tabular}{|l|l|l|} \hline
GDML tag 		&  Child tags				& pyg4ometry class 				\\ \hline
define 			& constant				& gdml.Constant				\\
				& variable					& gdml.Variable					\\
				& quantity					& gdml.Quantity				\\
				& expression 				& gdml.Expression				\\
				& position					& gdml.Position					\\
				& rotation					& gdml.Rotation				\\
				& scale					& gdml.Scale					\\
				& matrix					& gdml.Matrix					\\
materials 			& element					& geant4.Element				\\
				& isotope					& geant4.Isotope				\\
				& material					& geant4.Material				\\
solid  			& box					& geant4.solid.Box				\\
	  			& tube					&							\\
	  			& cutTube					&							\\
	  			& cone					&							\\
	  			& para					&							\\
	  			& trd						&							\\
	  			& trap					&							\\
	  			& sphere					&							\\
	  			& orb					&							\\
	  			& torus					&							\\
	  			& polycone				&							\\
	  			& genericPolycone			&							\\
	  			& polyhedra				&							\\
	  			& eltube					&							\\
	  			& ellipsoid					&							\\
	  			& elcone					&							\\
				& paraboloid				&							\\
				& hype					& 							\\
				& tet						& 							\\
				& xtru					& 							\\
				& twistedbox				& 							\\
				& twistedtrap				& 							\\
				& twistedtrd				& 							\\
				& twistedtube				& 							\\
				& arb8					& 							\\
				& tesselated				& 							\\
				& union					& 							\\
				& subtraction				& 							\\
				& intersection				& 							\\
				& multiUnion				& 							\\
				& scaledSolid				&							\\				
	  			& opticalsurface			&							\\
structure			& volume					& geant4.LogicalVolume			\\
				& assembly				& geant4.AssemblyVolume 		\\
				& physvol					& geant4.PhysicalVolume			\\ 
				& replicavol				& geant4.ReplicaVolume			\\ 
				& paravol					& geant4.ParameterisedVolume	\\ 
				& divisionvol				& geant4.DivisionVolume			\\
				& skinsurface				& geant4.SkinSurface			\\ 
				& bordersurface			& geant4.BorderSurface			\\ \hline

@@

\subsection{Overlap detection}
All PT transport codes can not handle spatial overlap between two geometric objects and likely are to have ill defined behaviour when tracking particles  
throughsuch a situation.  A key feature of pyg4ometry is the detection of potential overlaps efficiently, it does this by performing a CSG intersection between solid instances 
and determining if the resulting mesh is empty. Generally the overlap detection algorithm proceeds as

@@

\section{Geant4 and FLUKA conversion}
FLUKA geometry is based upon a limited set of primitive volumes which can be combined using logical operations. Many FLUKA primitives are infinite in extent, so 
to emulate FLUKA geometry construction large but finite extent Geant4 primitive solids are used. Table XXX shows the mapping between FLUKA and Geant4 objects.
An ANTLR based expression parser is used to load and comprehend the FLUKA input syntax, an example is shown in XXX. Each FLUKA body is matched to a Geant4 
solid and a FLUKA region is represented by a Geant4 CSG tree is formed using the pyg4ometry classes.

\begin{table}[hbt!]
\centering
\begin{tabular}{|l|l|l|} \hline
Fluka body			& note			& pyg4ometry class 				\\ \hline
RPP (done)			&				&							\\
BOX					&				&							\\
SPH (done)			&				&							\\
RCC (done)			&				&							\\
REC (done)			&				&							\\
TRC (done)			&				&							\\
ELL 					&				&							\\
WED					&				&							\\
RAW					&				&							\\
ARB					&				&							\\
XYP (done)			&				&							\\
XZP (done)			&				&							\\
YZP (done)			&				&							\\
PLA (done)			&				&							\\
XCC (done)			&				&							\\
YCC (done)			&				&							\\
ZCC (done)			&				&							\\
XEC (done)			&				&							\\
YEC (done)			&				&							\\
ZEC (done)			&				&							\\
QUA (done)			&				&							\\ \hline
\end{tabular}
\caption{Fluka geometry directives pyg4ometry classes.}
\end{table}

\begin{enumerate} 
\item infinite planes
\item disjoint unions 
\end{enumerate}

@@

\subsection{Editing existing CAD models}
Often existing CAD models have either missing details or too much detail.

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

@@

There are existing codes that have functionality similar to pyg4ometry. Tools exist to convert CAD files to GDML, open source examples are GUIMesh \cite{GUIMesh_Pinto}, 
commercial products include ESABASE2 \cite{ESABASE2} and FASTRAD \cite{FASTRAD}. 
CAD software can export data to STL. The existent set of software does not provide a complete set of tools to efficiently create complex geometries.

@@

\subsection{Geant Detector Mark-up Language} 
To exchange geometry descriptions Geant Detector Mark-up Language (GDML) was developed. Geant4 and ROOT can read and write GDML format and it is commonly
used as an exchange format for Geant4 geometries. 
\begin{table}[hbt!]
\centering
\begin{tabular}{|l|l|l|} \hline
GDML tag 		& Child tags				& pyg4ometry class 				\\ \hline
Define 			& Constant				& gdml.Constant				\\
				& Variable					& gdml.Variable					\\
				& Quantity					& gdml.Quantity				\\
				& Expression 				& gdml.Expression				\\
				& Position					& gdml.Position					\\
				& Rotation					& gdml.Rotation				\\
				& Scale					& gdml.Scale					\\
				& Matrix					& gdml.Matrix					\\ \hline
Materials 			& Element					& geant4.Element				\\
				& Isotope					& geant4.Isotope				\\
				& Material					& geant4.Material				\\ \hline
Solid  			& Box					& geant4.solid.Box				\\
	  			& Tube					& geant4.solid.Tubs				\\
	  			& CutTube				& geant4.solid.CutTubs			\\
	  			& Cone					& geant4.solid.Cons				\\
	  			& Para					& geant4.solid.Para				\\
	  			& Trd					& geant4.solid.Trd				\\
	  			& Trap					& geant4.solid.Trap				\\
	  			& Sphere					& geant4.solid.Sphere			\\
	  			& Orb					& geant4.solid.Orb				\\
	  			& Torus					& geant4.solid.Torus				\\
	  			& Polycone				& geant4.solid.Polycone			\\
	  			& GenericPolycone			& geant4.solid.GenericPolycone	\\
	  			& Polyhedra				& geant4.solid.Polyhedra			\\
	  			& Eltube					& geant4.solid.EllipticalTube		\\
	  			& Ellipsoid					& geant4.solid.Ellipsoid			\\
	  			& Elcone					& geant4.solid.EllipticalCone		\\
				& Paraboloid				& geant4.solid.Paraboloid			\\
				& Hype					& geant4.solid.Hyperboloid		\\
				& Tet						& geant4.solid.Tet				\\
				& Xtru					& geant4.solid.ExtrudedSolid		\\
				& TwistedBox				& geant4.solid.TwistedBox		\\
				& TwistedTtap				& geant4.solid.TwistedTrap		\\
				& TwistedTrd				& geant4.solid.TwistedTrd 		\\
				& TwistedTube				& geant4.solid.TwistedTube		\\
				& Arb8					& geant4.solid.GenericTrap		\\
				& Tessellated				& geant4.solid.TessellatedSolid 	\\
				& Union					& geant4.solid.Union				\\
				& Subtraction				& geant4.solid.Subtraction			\\
				& Intersection				& geant4.solid.Intersection 		\\
				& MultiUnion				& geant4.solid.MultiUnion 			\\
				& ScaledSolid				& geant4.solid.ScaledSolid		\\				
	  			& OpticalSurface			& geant4.solid.OpticalSurface		\\ \hline
Structure			& Volume					& geant4.LogicalVolume			\\
				& Assembly				& geant4.AssemblyVolume 		\\
				& PhysVol					& geant4.PhysicalVolume			\\ 
				& ReplicaVol				& geant4.ReplicaVolume			\\ 
				& ParaVol					& geant4.ParameterisedVolume	\\ 
				& DivisionVol				& geant4.DivisionVolume			\\
				& SkinSurface				& geant4.SkinSurface			\\ 
				& BorderSurface			& geant4.BorderSurface			\\ \hline

@@

\subsection{Overlap detection}
All PT transport codes can not handle spatial overlap between two geometric objects and likely are to have ill defined behaviour when tracking particles  
through such a situation.  A key feature of pyg4ometry is the detection of potential overlaps efficiently, it does this by performing a CSG intersection between solid instances 
and determining if the resulting mesh is empty. Generally the overlap detection algorithm proceeds as

@@

\section{FLUKA to Geant4 conversion}
FLUKA geometry is based upon a limited set of primitive volumes which can be combined using logical operations. Many FLUKA primitives are infinite in extent, so 
to emulate FLUKA geometry construction large but finite extent Geant4 primitive solids are used. Table XXX shows the mapping between FLUKA and Geant4 objects.
An ANTLR based expression parser is used to load and comprehend the FLUKA input syntax, an example is shown in XXX. Each FLUKA body is matched to a Geant4 
solid and a FLUKA region is represented by a Geant4 CSG tree is formed using the pyg4ometry classes.

\begin{table}[hbt!]
\centering
\begin{tabular}{|l|l|} \hline
FLUKA body							& pyg4ometry class 		\\ \hline
RPP 	(Rectangular parallelepiped)			& Box				\\
BOX	(General rectangular parallelepiped)		& Box				\\
SPH 	(Sphere)							& Orb				\\
RCC (Right circular cylinder)				& Cone				\\
REC (Right elliptical cylinder)				& EllipticalTube			\\
TRC (Truncated Right Angle Cone)			& Cone				\\
ELL (Ellipsoid of Revolution) 				& Ellipsoid				\\
WED/RAW (Right Angle Wedge)			& ExtrudedSolid		\\
ARB	(Arbitrary Convex Polyhedron)			& TessellatedSolid		\\
XYP 	($X$-$Y$ Infinite half-space)			& Box				\\
XZP 	($X$-$Z$ Infinite half-space)			& Box				\\
YZP 	($Y$-$Z$ Infinite half-space)			& Box				\\
PLA (Generic infinite half-space)			& Box				\\
XCC ($X$-axis Infinite Circular Cylinder)		& Tubs				\\
YCC ($Y$-axis Infinite Circular Cylinder)		& Tubs				\\
ZCC 	($Z$-axis Infinite Circular Cylinder)		& Tubs				\\
XEC 	($X$-axis Infinite Elliptical Cylinder)		& EllipticalTube			\\
YEC 	($Y$-axis Infinite Elliptical Cylinder)		& EllipticalTube			\\
ZEC ($Z$-axis Infinite Elliptical Cylinder)		& EllipticalTube			\\
QUA (Quadric surface) 					& TessellatedSolid		\\ \hline
\end{tabular}
\caption{Fluka geometry directives pyg4ometry classes.}
\end{table}

\begin{enumerate} 
\item length safety
\item infinite planes
\item disjoint unions 
\end{enumerate}

\section{Geant4 to FLUKA conversion}

\begin{table}[hbt!]
\centering
\begin{tabular}{| l | l | } \hline
Geant4 solid			& FLUKA constuction			\\ \hline
Box					& XYP, XZP, YZP 				\\
Tube					& ZCC, XYP	 				\\
CutTube				& ZCC, PLA					\\
Cone				& TRC						\\
Para					& ARB						\\
Trd					& ARB						\\
Trap					& ARB						\\
Sphere				& SPH, WED					\\
Orb					& SPH						\\
Torus				& (PLA)						\\
Polycone				& (PLA)						\\
GenericPolycone		& (PLA)						\\
Polyhedra				& (PLA)						\\
Eltube				& ZEC, XYP					\\
Ellipsoid				& ELL						\\
Elcone				& QUA						\\
Paraboloid			& QUA						\\
Hype					& QUA						\\
Tet					& PLA						\\
Xtru					& PLA						\\
TwistedBox			& (PLA)						\\
TwistedTtap			& (PLA)						\\
TwistedTrd			& (PLA)				 		\\
TwistedTube			& (PLA)						\\
Arb8					& ARB						\\
Tessellated			& (PLA)					 	\\
Union				& | zone1 | zone2				\\
Subtraction			& +zone1 -zone2				\\
Intersection			& +zone1 +zone2				\\
MultiUnion			& | zone1 | zone2 | zone3 | ...	\\
ScaledSolid			& geant4.solid.ScaledSolid		\\ \hline				
\end{tabular}
\end{table}

@@

\subsection{Editing existing models}
Often existing CAD models have either missing details or too much detail.
